### 2016-07-02 SIG Proceedings Template-Jan2015 Zip/paper.tex
write paper && modify code

% This is "sig-alternate.tex" V2.1 April 2013
% This file should be compiled with V2.5 of "sig-alternate.cls" May 2012
%
% This example file demonstrates the use of the 'sig-alternate.cls'
% V2.5 LaTeX2e document class file. It is for those submitting
% articles to ACM Conference Proceedings WHO DO NOT WISH TO
% STRICTLY ADHERE TO THE SIGS (PUBS-BOARD-ENDORSED) STYLE.
% The 'sig-alternate.cls' file will produce a similar-looking,
% albeit, 'tighter' paper resulting in, invariably, fewer pages.
%
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V2.5) produces:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) NO page numbers
%
% as against the acm_proc_article-sp.cls file which
% DOES NOT produce 1) thru' 3) above.
%
% Using 'sig-alternate.cls' you have control, however, from within
% the source .tex file, over both the CopyrightYear
% (defaulted to 200X) and the ACM Copyright Data
% (defaulted to X-XXXXX-XX-X/XX/XX).
% e.g.
% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the copyright line.
%
% ---------------------------------------------------------------------------------------------------------------
% This .tex source is an example which *does* use
% the .bib file (from which the .bbl file % is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission, you *NEED* to 'insert'
% your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% ================= IF YOU HAVE QUESTIONS =======================
% Questions regarding the SIGS styles, SIGS policies and
% procedures, Conferences etc. should be sent to
% Adrienne Griscti (griscti@acm.org)
%
% Technical questions _only_ to
% Gerald Murray (murray@hq.acm.org)
% ===============================================================
%
% For tracking purposes - this is V2.0 - May 2012

\documentclass{sig-alternate-05-2015}

\usepackage{algorithm}
\usepackage{algorithmic}

\begin{document}

% Copyright
\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\doi{10.475/123_4}

% ISBN
\isbn{123-4567-24-567/08/06}

%Conference
\conferenceinfo{PLDI '13}{June 16--19, 2013, Seattle, WA, USA}

\acmPrice{\$15.00}

%
% --- Author Metadata here ---
\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{A Failure Detection Approach of Data Center Networks\titlenote{(Produces the permission block, and
copyright information). For use with
SIG-ALTERNATE.CLS. Supported by ACM.}}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

%\numberofauthors{8} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\numberofauthors{2}
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Kai Shen\titlenote{Dr.~Trovato insisted his name be first.}\\
       \affaddr{School of Software}\\
       \affaddr{Shanghai Jiao Tong University}\\
       \affaddr{Shanghai, China}\\
       \email{knshen@sjtu.edu.cn}
% 2nd. author
\alignauthor
????\titlenote{Dr.~Trovato insisted his name be first.}\\
       \affaddr{School of Software}\\
       \affaddr{Shanghai Jiao Tong University}\\
       \affaddr{Shanghai, China}\\
       \email{????@sjtu.edu.cn}
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.
\maketitle

\begin{abstract}
This paper provides a sample of a \LaTeX\ document which conforms,
somewhat loosely, to the formatting guidelines for
ACM SIG Proceedings. It is an {\em alternate} style which produces
a {\em tighter-looking} paper and was designed in response to
concerns expressed, by authors, over page-budgets.
It complements the document \textit{Author's (Alternate) Guide to
Preparing ACM SIG Proceedings Using \LaTeX$2_\epsilon$\ and Bib\TeX}.
This source file has been written with the intention of being
compiled under \LaTeX$2_\epsilon$\ and BibTeX.

The developers have tried to include every imaginable sort
of ``bells and whistles", such as a subtitle, footnotes on
title, subtitle and authors, as well as in the text, and
every optional component (e.g. Acknowledgments, Additional
Authors, Appendices), not to mention examples of
equations, theorems, tables and figures.

To make best use of this sample document, run it through \LaTeX\
and BibTeX, and compare this source code with the printed
output produced by the dvi file. A compiled PDF version
is available on the web page to help you with the
`look and feel'.
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below.
%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Embedded systems}
\ccsdesc[300]{Computer systems organization~Redundancy}
\ccsdesc{Computer systems organization~Robotics}
\ccsdesc[100]{Networks~Network reliability}

%
% End generated code
%

%
%  Use this command to print the description
%
\printccsdesc

% We no longer use \terms command
%\terms{Theory}

\keywords{Failure detection; Quality of Service; Data center networks}

\section{Introduction}
Data center.

The rest of the paper is organized as follows. Section 2 discusses related work. Section 3 presents data center networks architecture. Section 4 provides an overview of our approach, while the details are described in Section 5. The experimental evaluation along with experiments results is presented in Section 6. Finally, Section 7 concludes this paper.

\section{Related Work}
In this section, we will present various relevant works including failure detectors, failure detection architecture in distributed systems and failure detection in data center networks.

\subsection{Failure Detectors}
A failure detector (FD) is widely recognized as an oracle to intelligently suspect failed processes \cite{chandra1996unreliable}. The monitored process periodically send heartbeat messages to its detector to prove its liveness.

Chen et al. proposed a FD that provides QoS \cite{chen2002quality}. It estimates the expected arrival time (EAs) of the next heartbeat message according to a slide window storing n most recent arrived messages. EAs determine the deadline that the detector will wait for the next heartbeat before suspecting the monitored process. \textit{Chen FD} can adjust current network condition and set the timeout threshold adaptively by referring to recent heartbeats. Bertier et al. proposed a similar FD \cite{bertier2002implementation}, whereas it uses a dynamic way to compute error margin of \textit{Chen FD}. Other FDs like \textit{$\phi$ Accrual FD} \cite{hayashibara2004varphi} and \textit{Satzger FD} \cite{satzger2007new} can output a continuous accrual failure value other than a binary value at any time which stands for the failure probability of the process.

\subsection{Failure Detection Architecture}
A failure detection architecture aims to provide the service of monitoring nodes in large scale distributed systems in a scalable way.
Roughly speaking, there are two kinds of architectures. The first one is hierarchical architecture \cite{felber1999failure}, all the nodes are partitioned into different groups, each group has a leader node. Within a group, the leader node is responsible for monitoring all the nodes. Leader nodes periodically send node status information of his group to other leader nodes. Hierarchical architecture can reduce the number of heartbeat messages effectively, but it exists \textit{Single Point of Failure} problem, when the leader node crashes, it needs to select a new leader.

Another alternative solution is Gossip-Style architecture \cite{van1998gossip}. Each node maintains a list containing the \textit{heartbeat counter} for each node in the system. Every T$_{gossip}$ seconds, each node firstly update \textit{heartbeat counter} of itself and then randomly select another node to send its list to. Upon receiving a message, the node will merge two lists and update each node's \textit{heartbeat counter} with the bigger one. If the \textit{heartbeat counter} does not update after T$_{fail}$ seconds, the node will be marked as crashed. If the \textit{heartbeat counter} does not update after T$_{cleanup}$ (T$_{cleanup}$
$\ge$T$_{fail}$) seconds, it will be removed from the list. This approach has very low bandwidth occupation because it only produces $n$ messages each epoch where $n$ is the size of the system. However, with the increasing of crashing nodes, it needs more time to detect these failures.

\subsection{Failure Detection in Networks}
Different from the work mentioned above, failure detection in networks should consider network topology and localize failure positions at IP layer. Gill et al. present the first large-scale survey on failures in data center networks \cite{gill2011understanding}.

Also, there exists a large body of work on applying statistical approaches to failure localization in networks. Sherlock \cite{bahl2007towards} localizes failures to limited network components by capturing dependencies between components. Shrink \cite{kandula2005shrink} leverages Bayesian Network model to diagnose root cause of IP network failures. Herodotou et al. \cite{herodotou2014scalable} actively inject ping traffic into network and analyze the ping results overlaid on top of the network topology. It will generate a ranked list containing the suspected failed device and links along with their failure score that can best explain the observed ping data.

\section{Data Center Network Architecture}
In this section, we will take a brief look at common data center network architecture.

A data center is a facility used to house computer systems and associated components, it usually contains hundreds of thousands devices and links. Modern data centers use hierarchical architecture. Figure 1 depicts a conventional data center network architecture \cite{gill2011understanding, greenberg2009vl2, herodotou2014scalable}. At the bottom layer are racks of servers, typically there are 20 to 40 servers in each rack, and each server in the same rack connects to a Top of Rack (ToR) Switch via a 1 Gbps link. Each ToR then connects to a primary and a backup Aggregation Switch (AS) for redundancy. ASs are connected to Access Routers (ARs) forming a complete bipartite graph. Finally, ARs are connected to Core Routers (CRs).

To limit overheads and to isolate different services, servers are partitioned into virtual LANs (VLANs). All links in the data center use Ethernet as link layer protocol and physical connections are a mix of copper and fiber cables. In addition to the devices in Figure 1, there may exist other devices like load balancers (LBs) and firewalls.

Modern data center usually employs \textit{Equal Cost MultiPath} (ECMP) routing within the data center for fault tolerance, which means that there are multiple paths from source to destination. Thus in this work, we will apply ECMP routing to the simulated network by enabling ECMP routing configurations in ns-3.

% figure
\begin{figure}
\centering
\includegraphics{topology}
\caption{Conventional Data Center Network Topology}
\end{figure}

\section{Approach Overview}
In this section, we will introduce the whole design of our failure detection approach.

Different network device should have different quality of service (QoS). Gill et al. \cite{gill2011understanding} have shown the diversity of different devices' failure characteristics. Unfortunately, none of the work in Section 2 treats these devices differently. We firstly divide network devices into two categories.

\begin{itemize}
\item \textbf{Imperative Device.} This category includes servers, ToRs and links between ToRs and servers. Typically servers run various user applications and store user data. The crash of servers may cause user-perceived faults. The crash of a ToR will make servers under it unreachable. Also Gill et al. declare that the downtime of ToRs may be very long. Thus the failure detection for these devices must be fast enough besides the accuracy requirement.

\item \textbf{Non-imperative Device.} It includes links, other routers and switches. A single failure of a link or a router does not affect the functionality of a data center because of path redundancy. So administrators care more about how effective the detect service is (The ability to identify actually happening failures).

\end{itemize}

% figure
\begin{figure}
\centering
\includegraphics{system}
\caption{Failure Detection System Architecture}
\end{figure}

Figure 2 illustrates the entire architecture of our failure detection system. It mainly contains two modules. The servers failure detection system combines centralized with decentralized structures. A server can both play the role of a detector and a monitored object. Each server is monitored by multiple FD processes installed on different servers for redundancy. A server can also monitor a couple of servers. Once the FD outputs the crash of a server, it will send a crash report message to \textbf{Crash Decider}. \textbf{Crash Decider} will determine the final status of this server.

For the detection of non-imperetative devices, \textbf{Traffic Collector} periodically collects traffic data of routers and switches and send telemetry data to \textbf{Traffic Analysis Worker}. \textbf{Traffic Analysis Worker} finally generate failure alerts by performing analysis on these data.

\section{Failure Detection Approach}
In this section, we will discuss the detailed design of different modules of the failure detection system respectively.

\subsection{K-detectors Model}
To detect the crash of servers and ToRs effectively and efficiently, this paper propose a detection model named $K$-detectors model. Our detection model is based on the assumption that once the server or ToR crashes, it never recovers by itself. The QoS of failure detection are as follows:

\begin{itemize}
\item \textbf{Efficiency.} It measures how fast the system can report the crash after it crashes. The system is more efficient if it takes less time to detect a failure.
\item \textbf{Completeness.} It measures whether the system can detect all the crashed objects. The system is more complete if it rules out fewer true crashed ones.
\item \textbf{Soundness.} In some scenarios, the behavior of the network may vary. For example, message loss and message delay may happen sometimes. The detection system should be sound enough to distinguish between such condition and a real crash.

\end{itemize}

In $K$-detectors model, each server is monitored by $K$ FDs on different servers where $K$ is a parameter tuned by users. When a FD considers a server crashes, it sends a crash report message to \textbf{Crash Decider}. If the output of the FD transforms from suspect to trust, it re-sends a message to revoke the most recent crash report.

\begin{equation}
S_{s,t}=\left\{
    \begin{array}{lcl}
    \text{suspect} & \text{if} & |CRM_{s,t}|\geq\left\lfloor(K+1)/2\right\rfloor \\

    \text{trust} && \text{otherwise}
    \end{array}
\right.
\end{equation}
where $CRM_{s,t}$ denotes the collection of crash report messages for server $s$ until time $t$, $S_{s,t}$ denotes the status of server $s$ at time $t$. If \textbf{Crash Decider} receives $\left\lfloor(K+1)/2\right\rfloor$ crash messages of a server, it considers the server to have crashed.

\subsection{Monitoring Rule Inference}
To form the $K$-detectors model and also to meet QoS to the greatest extent, we propose the following specifications of monitoring relationships among servers:
\begin{enumerate}
    \item Each sever monitors $K$ servers. This balances monitor intensity of each server.
    \item A server cannot monitor itself.
    \item A server is monitored by $K$ different servers.
    \item $K-1$ of the $K$ detectors are on the servers at the same rack.
    \item The monitoring rule should ensure each link covered by heartbeat messages.
    \item Each rack monitors servers of other racks which vary from different racks averagely.
\end{enumerate}

\subsubsection*{Rationale}
\quad Intuitively, the use of $K$ FDs is able to adapt to bursty-traffic conditions like message loss because the determination of a failure requires at least $\left\lfloor(K+1)/2\right\rfloor$ FDs' agreement. Also, $K$-detectors can prevent crashes of FDs to some extent. Theoretically, a failure can still be detected even when $K-\left\lfloor(K+1)/2\right\rfloor=\left\lfloor K/2\right\rfloor$ of its detectors crashed.

On the other hand, the fourth specification above can make the best of locality of LAN, which reduces the probability of degeneration on detection time in turn.

Specification five is the presupposition of link failure detection. It is easy to see that, some links cannot be distinguished from faulty links without this specification.

The last one is an extension of specification five. It is for the consideration of ToRs' detection, we will discuss it in detail in Section 5.4.

\subsection{FD Algorithm}
FD algorithm is used to decide the status of one monitored object by one FD. Our FD algorithm is almost the same as \textit{Chen} FD \cite{chen2002quality} except for some minor changes.

\begin{equation}
EA_{l+1} \approx \frac{1}{n}(\sum_{i=1}^{n}A_i - \eta s_i)+(l+1)\eta
\end{equation}
this formula estimates the expected arriving time of next message, where $s_i$ is the sequence number of heartbeat message, $A_i$ is the receipt time of message $i$, $\eta$ is message send interval. Obviously, this formula normalizes each $A_i$ by shifting it backward $\eta s_i$ time units.

After that, we utilize hyperbolic tangent function to model current suspicious level:

\begin{equation}
\tau(t_{now}) = |tanh(t_{now}-EA_{l+1})|
\end{equation}
suspicious level $\tau(t_{now})$ always ranges from 0 to 1, when $t_{now}$ grows bigger, the suspicious level is closer to 1, which implies a higher failure probability. A continuous valued rather than binary valued output decouples the interpretation of failure data from the failure monitoring mechanism, user applications can set their own threshold according to their demands.

\subsection{Failure Detection for ToRs}
Apart from determination of servers' crashes, \textbf{Crash Decider} is also responsible for failure detection for ToRs. If most or even all servers in a rack is reported crash by FDs of other racks, it is more likely that the top of rack switch crashes rather than the servers.

\begin{equation}
S_{ToR,t}=\left\{
    \begin{array}{lcl}
    \text{suspect} & \text{if} & |\left\{s|s\in ToR\right\} \cap \left\{s|S_{s,t}=\text{suspect}\right\}| \\
    && \geq \alpha \cdot m\\
    \text{trust} && \text{otherwise}
    \end{array}
\right.
\end{equation}
where $m$ is the number of servers per rack, $\alpha$ is a  proportionality coefficient measuring how strict the criterion is to determine failure of a ToR. When \textbf{Crash Decider} gets more than $\alpha \cdot m$ crash report messages of servers under the ToR crash, it will report the crash of ToR. The last specification in Section 5.2 ensures that once a ToR is down, each other rack has the chance to find this failure. An extreme counter-example is that if the outside detectors of rack 1 servers are all in rack 2, once ToR of rack 2 is down, \textbf{Crash Decider} never perceives the status of ToR of rack 1.

To sum up, Algorithm 1 illustrates the pseudo-code for overall detection algorithm of \textbf{Crash Decider}.

\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\begin{algorithm}
	\caption{Detection Algorithm of \textbf{Crash Decider}}
	%\label{alg0}
	\begin{algorithmic}[1]
		\REQUIRE $K$, $\eta$, $n$, $\alpha$, $m$, $\tau_m$
		\ENSURE $S_{i,t}$, $i$ is an imperative device

        \FOR{$i \text{ in imperative devices}$}
            \STATE $S_{i,t_0} \leftarrow trust$
        \ENDFOR
	    \WHILE{$True$}
            \STATE $List<Integer> failsEachRack \leftarrow \emptyset$
            \FOR{each $i$ in Servers}
                \STATE $rackID \leftarrow getRackID(i)$
                \IF{$|CRM_{i,t_{now}}| \geq \left\lfloor(K+1)/2\right\rfloor$}
                    \STATE $S_{i,t_{now}} \leftarrow suspect$
                \ELSE
                    \STATE $S_{i,t_{now}} \leftarrow trust$
                    \IF{$\text{only one CRM coming from other rack}$}
                    \STATE $failsEachRack.get(rackID)\text{++}$
                    \ENDIF
                \ENDIF
            \ENDFOR

            \FOR{each i in ToRs}
                \IF{$failsEachRack.get(i) \geq \alpha \cdot m$}
                    \STATE $S_{i,t_{now}} \leftarrow suspect$
                \ELSE
                    \STATE $S_{i,t_{now}} \leftarrow trust$
                \ENDIF
            \ENDFOR
		\ENDWHILE
	\end{algorithmic}
\end{algorithm}
It firstly determines the status of each server (line 6 to 16), and then determines the status of ToRs (line 17 to 23). When detecting server failures, \textbf{Crash Decider} also records necessary information to find potential ToR crashes (line 12 to 13).

\subsection{Failure Detection for Links}
The failure detection for links is another important part of our detection system, this paper presents the following QoS:
\begin{itemize}
  \item \textbf{Precision.} Precision indicates how many suspected links are true failed.
  \item \textbf{Completeness.} This measures whether our approach can find all the faulty links.
  \item \textbf{Discrimination.} Because we will provide a short ranked list of suspected objects, this measures to what degree can our approach distinguish between normal links and faulty links.
\end{itemize}

The key idea of our approach is to capture network traffic changes over time. If the network traffic changes significantly compared with traffic in normal conditions, we believe that there exists failures somewhere in the network. The insight is that network traffic distribution is associated with routing protocol and monitoring relationships of servers, the drastic changes of traffic probably imply failures. For example, if the network topology is just as Figure 1 shows, when the link between the leftmost AR and the leftmost AS fails, the traffic on the leftmost AR will decrease and the rightmost AR's traffic will increase according to ECMP routing.

To collect required traffic data, we simply install \textit{tcpdump} on each router and switch. These data will be collected to \textbf{Traffic Analysis Worker} for further analysis.

In order to model this data analysis problem, this paper proposes a Back Propagation Neural Networks (BPNN) based method to compute failure probability of each link.

\subsubsection*{Feature Extraction}
\quad The first step is to extract rational features from raw traffic data. To address this issue, we extract $v \cdot n^2$ features in total. $v$ is number of devices (CRs, ARs, ASs), each device has $n^2$ features where $n$ is the number of racks. In particular, each feature denotes traffic of one direction. Obviously, there are $n$ choices of sources and $n$ choices of destinations, so the total number of possible traffic directions is $n^2$.
Formally, the traffic features during a time period $\Delta T$ is presented as \boldmath $x$
\unboldmath $=($\boldmath $x_1, x_2, ..., x_v$\unboldmath $)$, \boldmath $x_i$
\unboldmath $=(x_i^1, x_i^2, ..., x_i^{n^2})$
where $x_i^j$ stands for direction $j$ traffic of device $i$ in $\Delta T$.

\subsubsection*{BPNN Model}
\quad Suppose the number of links is $e$, for each time period $\Delta T$, the output of our model should be
\boldmath $y$
\unboldmath $=(y_1, y_2, ..., y_e)$, where $y_i$ denotes failure probability of link $i$. We apply a BPNN model to localize the faulty link during $\Delta T$. As is shown in Figure 3, BPNN is a multi-layer perceptron architecture:
\begin{figure}
  \centering
  \includegraphics{ANN}
  \caption{Multi-layer Perceptron Architecture}
\end{figure}
To reduce the computational work of training, we only set one hidden layer, both the hidden layer neurons and output layer neurons utilize Sigmoid function as the activation function. Take the output layer as an example:
$y_i = Sigmoid(\beta_i - \theta_i)$ where $\theta_i$ is a constant term of the perceptron.

\subsubsection*{Training of BPNN}
\quad To train the neural network, we apply the classical error BackPropagation (BP) algorithm \cite{rumelhart1988learning}. BP algorithm is based on gradient descent strategy.% write BP algorithm?

To get train data, we manually disable links to simulate real link failures. To be specific, each time we disable one link for some time. After that, we collect dump data from each router and switch, extracting direction features just as mentioned above. The next step is to split the traffic data into different chunks, with each chunk representing the traffic data within one time unit ($\Delta T$). Each traffic chunk forms one train instance.

\subsubsection*{The Detection Algorithm}
\quad BPNN model is only able to identify the faulty link during one time unit, in the following part, we will introduce how we can detect multiple faults over time.

First of all, we assume that there are at most one link failure during a time unit ($\Delta T$). This assumption is partially supported by the finding proposed by Gill et al. They find that on an average, link failures tend to be separated by a period of about one week \cite{gill2011understanding}.

\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\begin{algorithm}
	\caption{\textbf{Traffic Analysis Worker} Algorithm}
	\begin{algorithmic}[1]
        \REQUIRE \boldmath $x^t$ \unboldmath$=($\boldmath $x_1^t, x_2^t, ..., x_v^t$\unboldmath $)$
		\ENSURE $F_i \text{ for } i \in E$

        \STATE $\text{divide traffic data }$\boldmath $x^t$$\text{ into }$\unboldmath $\left\lfloor t/\Delta T \right\rfloor \text{ chunks}$

        \FOR{$i \in E$}
            \STATE $F_i \leftarrow 0$
        \ENDFOR
        \FOR{$\text{each data chunk } $ \boldmath $x$}
        \STATE \boldmath $y$\unboldmath $\leftarrow BPNN($\boldmath $x$\unboldmath $)$
            \FOR{$i \in E$}
                \STATE $F_i \leftarrow F_i + y_i$
            \ENDFOR
        \ENDFOR

        \FOR{\unboldmath $i \in E$}
            \STATE $F_i \leftarrow \frac{\Delta T F_i}{t}$
        \ENDFOR
        \STATE $\text{sort }$\boldmath $F$\unboldmath$\text{ in descending order}$
	\end{algorithmic}
\end{algorithm}

Algorithm 2 shows the algorithm of \textbf{Traffic Analysis Worker}. The input of the algorithm is the refined time series traffic data over time $T$. Firstly, traffic data is divided into chunks. Each chunk represents a time unit $\Delta T$ (line 1) of traffic. The last chunk will be dropped if its time period is less than $\Delta T$. For each of traffic data chunk, we utilize the trained BPNN model to compute failure score of links during this time unit. The final failure score over $t$ is computed as the arithmetic mean of failure scores of all time units (line 5 to 13). This solution inevitably benefits the early happened faulty links. For example, when there are only one faulty links in the network, the failure score of the sole faulty link will be very high, which in turn makes its final failure score higher. However, it does not matter because we care more about whether it can find as many failures rather than the ranking.

\section{Evaluation}
In this section, we will present how we evaluate our approach in terms of efficiency and effectiveness. Experiments results will also be shown in this section.

\subsection{Experimental Settings}
To build up a data center network, we deploy ns-3 on PC. ns-3 is a discrete-event network simulator for Internet systems, targeted primarily for research and educational use \cite{henderson2008network}. we install ns-3 on single PC with 2 cores, 3.4 GHz CPU and 4GB RAM. The version of ns-3 we use is ns-3.24.

We set up a simulated data center network following the topology in Figure 1. There are a total of 120 servers (20 servers $\times$ 6 racks), 14 routers and switches and 140 links. For each link, data rate is 100 Mbps, delay is 100 ms.

\subsection{Research Questions}
The objective of our experiments is to validate the effectiveness and efficiency of the proposed approach compared with baselines. For this purpose, we set up the following research questions:

\subsubsection*{RQ1: How about the QoS of K-detectors model?}
\quad RQ1 evaluates QoS of failure detection for servers. To answer this question, we firstly set four metrics \cite{chen2002quality} corresponding to the QoS in Section 5.1.
\begin{itemize}
  \item \textbf{Average Detection Time ($AT_D$)} is the average time that elapses from the crash of a server until the crash is reported.
  \begin{equation}
    AT_D = \frac{\sum_i(T_{D,i} - T_{C,i})}{n}
  \end{equation}
  where $n$ is the number of detected crashes, $T_{D,i}$ is the detection time of $i$th crash $i$, $T_{C,i}$ is the crash time of $i$th crash.

  \item \textbf{Average Mistake Rate ($AR_M$)} measures the average rate at which a FD suspects a correct object. It equals the number of such mistakes per unit of time. Suppose Figure 4 depicts the monitoring of the object in one time unit, the mistake rate of this FD is two.
\begin{figure}
  \centering
  \includegraphics{metric}
  \caption{Average Mistake Rate}
\end{figure}

  \item \textbf{Query Accuracy Probability ($P_A$)} is the probability that the detection system outputs the correct results when queried at a random time.

  \item \textbf{Number of Undetected Crashes} measures the number of ignored real crash servers.
\end{itemize}

We compare our work with the traditional hierarchical architecture \cite{felber1999failure} in terms of the metrics above.

\subsubsection*{RQ2: Can K-detectors model benefit ToR crash detection?}
\quad The evaluation of failure detection for ToRs is almost the same as servers. We will evaluate the impact of different values of $\alpha$.

\subsubsection*{RQ3: How about the QoS of BPNN based model?}
\quad This RQ evaluates whether the proposed approach is more effective in detecting link failures. We make another three metrics:
\begin{itemize}
    \item \textbf{Precision@k.} We make the use of this metric in order to evaluate to what degree the detection is accurate. This metric is appropriate because it does not take into account the positions in the list which caters to our requirements \cite{mogotsi2010christopher}.
    \begin{equation}
        precision@k = \frac{1}{n}\sum_{i=1}^n \frac{fail_{i,k}}{k}
    \end{equation}
    where $fail_{i,k}$ denotes the number of true failures for failure event $i$ in the top $k$ returned results. $n$ is the number of failure events, $k$ equals to the number of true failures. A perfect result is all the $k$ failures are ranked at the top $k$ positions.

    \item \textbf{MRR.} We also make use of Mean Reciprocal Rank (MRR), which is a statistic measure for evaluating a process that produces a ranked list to a sample of queries, the list contains items ordered by probabilities \cite{radev2002evaluating}. It is the average multiplicative inverse of the rank of the first faulty link:
        \begin{equation}
            MRR = \frac{1}{n}\sum_{i=1}^n \frac{1}{rank_i}
        \end{equation}

    \item \textbf{Discriminative Significance} (DS). DS is based on the notion of entropy. Entropy is a measure of unpredictability of information content \cite{shannon2001mathematical}.
        \begin{equation}
            DS = -\sum_{i=1}^e F_i \cdot log_2 F_i
        \end{equation}
        where $F_i$ is a normalized failure score of link $i$. A smaller DS implies that the results of ranked list is more determinate, which makes it easier for further troubleshooting.

\end{itemize}

We will compare this part of our approach with the work of Herodotou et al. \cite{herodotou2014scalable}.

\subsection{Experiments Results}

\subsubsection*{RQ1: QoS of K-detectors Model}
\quad We evaluate our approach in the presence of different degrees of message loss and server crashes. To be fair, FD parameters of all the methods are set to the same: $\tau_m=0.001$, slide window size $n=100$. Heartbeat message send interval $\eta$ is 100 ms. The whole application is run for 1 min.

\begin{figure}
  \centering
  \begin{tabular}{c}
  \includegraphics[scale=0.7]{r1} \\
  \includegraphics[scale=0.7]{r2}
  \end{tabular}
  \caption{Effect of Message Loss on $P_A$ and $AM_R$}
\end{figure}
Figure 5 depicts $P_A$ and $AR_M$ of K-detectors and hierarchical architecture under different number of message loss without server crashes. Both the time and channel of message loss is chosen randomly. When loss rate is under 1\%, 5-detectors has about an 8.2\% improvement in $P_A$ compared with hierarchical architecture. When loss rate is over 0.6\%, 5-detectors makes more mistakes than others, this is probably because other methods make mistakes with relatively longer time. In a nutshell, with a larger parameter $K$, our approach behaves better in the presence of normal message loss. However, it cannot maintain with the increasing number of message loss.

\begin{figure}
  \centering
  \begin{tabular}{c}
  \includegraphics[scale=0.7]{r3} \\
  \includegraphics[scale=0.7]{r4} \\
  \includegraphics[scale=0.7]{r5} \\
  \end{tabular}
  \caption{Effect of Crashes on $P_A$, $AR_M$ and $AT_D$}
\end{figure}
Subsequently, we investigate it when there exists server crashes. We set message loss rate to 0.1\%, both the time and server ID of crashes are selected randomly. As is shown in Figure 6, 5-detectors has a better $P_A$ and $AR_M$ than others in most cases. In terms of $AD_T$, a smaller $K$ results in a quicker detection because it needs fewer FDs to reach an agreement. In general, $AD_T$ of 5-detectors has a 22.4\% increase, while 2-detectors only has a 7.3\% increase compared with hierarchical architecture. Figure 7 shows the number of undetected crashes, hierarchical architecture tends to behave worse than K-detectors, which reveals the \textit{Single Point of Failure} problem of hierarchical architecture.

\begin{figure}
  \centering
  \includegraphics[scale=0.7]{r6}
  \caption{Number of Undetected Crash Servers}
\end{figure}

\subsubsection*{RQ2: ToR crash Detection}
\quad To answer this question,

The drawback of our approach is that once occurring ToR failures, we cannot detect server crashes within the rack before the ToR recovering.

\subsubsection*{RQ3: QoS of BPNN Based Model?}
\quad To answer this question, we wants to firstly explain why we select BPNN rather other popular classifiers. We select 3 other algorithms including Decision Tree (C4.5) \cite{quinlan2014c4}, Bayesian Networks (BN) \cite{jensen1996introduction} and Support Vector Machine (SVM) \cite{scholkopf1999advances}. We apply these algorithms to train data by 10-fold cross-validation. Results are all shown in Table 1. Obviously BPNN outperforms other algorithms.

\begin{table}
    \centering
    \caption{Results of Different Classifiers}
    \begin{tabular}{|c|c|c|c|} \hline
        \textbf{Classifier} & \textbf{Precision} & \textbf{Recall} & \textbf{F-Measure} \\ \hline
        \hline
        \textbf{BPNN} & 98.8\% & 98.8\% & 98.8\% \\ \hline
        \textbf{C4.5} & 96.4\% & 96.3\% & 96.3\% \\ \hline
        \textbf{BN} & 96.9\% & 96.6\% & 96.7\% \\ \hline
        \textbf{SVM} & 96.9\% & 96.6\% & 96.7\% \\ \hline
    \end{tabular}
\end{table}

\begin{figure}
  \centering
  \includegraphics[scale=0.7]{precision} \\
  \caption{Precision@k of Different Failure Group Size}
\end{figure}

Empirically, parameters for BPNN is set as follows: learning rate equals 0.3, number of iterations equals 500, momentum equals 0.2, number of neurons at the hidden layer is $(\# attributes + \# classes) / 2$, $\Delta T = 1 $ s. The selection of $\Delta T$ is a tradeoff, when $\Delta T$ grows bigger, it may contain multiple failures which violates our assumption, while a smaller $\Delta T$ may add more occasionality to traffic data.

\begin{figure}
  \centering
  \includegraphics[scale=0.7]{mrr} \\
  \caption{MRR of Different Failure Group Size}
\end{figure}

We manually make some link failures by disabling links in ns-3, we make 5 random failure events for different size of failure group respectively. Both the link ID and failure time are selected randomly. As Gill et al. finds that link failures tend to be isolated \cite{gill2011understanding}. Only 41\% of the link failure groups contain more than one faults, only 10\% contain more than four faults. Consequently we only make failure events whose failure group size is no more than four.

\begin{figure}
  \centering
  \includegraphics[scale=0.7]{ds} \\
  \caption{DS of Different Failure Group Size}
\end{figure}

Figure 8,9 and 10 show the results of out experiments. Our approach has a better performance in terms of percision@k and MRR. In particular, there is an at least 73.3\% precision@k performance, which indicates 73.3\% of the failures can be found at the most top of the suspects averagely.

On the other hand, there is a huge improvement in terms of DS. The approach proposed by Herodotou et al. tends to assign a wider range of failure scores to multiple links, while our approach is better at finding the relevant links. It mainly stems from the fact that our approach is based on a multi-layer perceptron architecture, it can precisely capture the characteristic of each link failure. But one of the drawbacks of our approach is that it requires much data and time to train BPNN offline to ensure its generalization ability, it cannot adapt to network topology changes as well.

\section{Conclusions}
In this paper, we described a complete failure detection approach for data center networks which leverages data mining techniques. 
%\end{document}  % This is where a 'short' article might terminate

\section{Acknowledgments}
This work is partially supported by the National High-Technology Research and Development Program of China under Grant No.2014AA01A301.

%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{paper}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%APPENDICES are optional
%\balancecolumns

\end{document}